---
title: "langevitour examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{langevitour examples}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<style>
h2,h3,h4 { margin-top: 200px; }
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

These are some example datasets visualized using [langevitour](https://github.com/pfh/langevitour).

Things to try:

* Drag labels onto the plot.
* Activate "point repulsion".
* Experiment with "damping" and "heat".

```{r install, eval=FALSE}
remotes::install_github("pfh/langevitour")
```

```{r setup}
library(langevitour)
```

## Olive oils

A dataset from the [tourr](http://ggobi.github.io/tourr/) package.

By default, langevitour does not scale variables individually. This dataset rather requires scaling, so we supply our desired scaling.

```{r}
olives <- tourr::olive[,3:10]
olives_scale <- apply(olives, 2, sd) * 6
olives_group <- paste(tourr::olive$region, tourr::olive$area)

langevitour(olives, group=olives_group, scale=olives_scale)
```


## Iris

The classic.

```{r}
langevitour(iris[,1:4], iris$Species)
```


## fake_trees

A dataset in the [liminal](https://sa-lee.github.io/liminal/) package, originally from PHATE.

* "Point repulsion" is especially effective in this dataset.

* You can manipulate it like a puppet by dragging the groups (0-9) onto the plot.


```{r}
langevitour(liminal::fake_trees[,1:100], liminal::fake_trees$branches)
```


## fake_trees with k-nearest neighbor smoothing

```{r}
X <- as.matrix(liminal::fake_trees[,1:100])
knn <- RANN::nn2(X, k=10)$nn.idx
X_smooth <- lapply(seq_len(nrow(X)), function(i) colMeans(X[knn[i,],]))
X_smooth <- do.call(rbind, X_smooth)

langevitour(X_smooth, liminal::fake_trees$branches)
```


## Hyper-cube

Some geometry from the [Geo Zoo](https://github.com/schloerke/geozoo).

```{r}
langevitour(geozoo::cube.dotline(5)$points)
```

## Torus

```{r}
langevitour(geozoo::torus.flat(p=4)$points)
```

## pdfsense

Another dataset in the [liminal](https://sa-lee.github.io/liminal/) package. Each row has been scaled to a unit vector.

```{r}
X <- as.matrix(liminal::pdfsense[,-(1:6)])
X <- X / sqrt(rowSums(X*X))
langevitour(X, liminal::pdfsense$Type)
```

## Stiefel manifold $V_a(\mathbb{R}^b)$

The orthonormal projections of b dimensions onto a.

Stiefel manifolds are what langevitour tours!

```{r}
stiefel <- function(a,b) {
    k <- 5000
    
    X <- matrix(nrow=k, ncol=a*b)
    colnames(X) <- paste0("X[",rep(seq_len(a),b),",",rep(seq_len(b), each=a),"]")
    for(i in seq_len(k)) {
        mat <- matrix(rnorm(a*b), nrow=a)
        decomp <- svd(mat)
        mat <- decomp$u %*% t(decomp$v)
        X[i,] <- as.vector(mat)
    }
    
    langevitour(X)
}

stiefel(2,3)
```

```{r}
stiefel(2,2)
```