---
title: "langevitour examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{langevitour examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
h2,h3,h4 { margin-top: 200px; }
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

These are some example datasets visualized using [langevitour](https://github.com/pfh/langevitour).

Things to try:

* Drag labels onto the plot.
* Activate "point repulsion".
* Experiment with "damping" and "heat".

```{r install, eval=FALSE}
remotes::install_github("pfh/langevitour")
```

```{r setup}
library(langevitour)
```

## Olive oils

A dataset from the [tourr](http://ggobi.github.io/tourr/) package.

```{r}
langevitour(scale(tourr::olive[,3:10]), paste(tourr::olive$region, tourr::olive$area))
```

## Iris

The classic.

```{r}
langevitour(iris[,1:4], iris$Species)
```

## fake_trees

A dataset in the [liminal](https://sa-lee.github.io/liminal/) package, originally from PHATE.

* "Point repulsion" is especially effective in this dataset.

* You can manipulate it like a puppet by dragging the groups (0-9) onto the plot.


```{r}
langevitour(liminal::fake_trees[,1:100], liminal::fake_trees$branches)
```

## Hyper-cube

Some geometry from the [Geo Zoo](https://github.com/schloerke/geozoo).

```{r}
langevitour(geozoo::cube.dotline(5)$points)
```

## Torus

```{r}
langevitour(geozoo::torus.flat(p=4)$points)
```

## pdfsense

Another dataset in the [liminal](https://sa-lee.github.io/liminal/) package. Each row has been scaled to a unit vector.

```{r}
X <- as.matrix(liminal::pdfsense[,-(1:6)])
X <- X / sqrt(rowMeans(X*X))
langevitour(X, liminal::pdfsense$Type)
```
