---
title: "Usage example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Usage example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<script>
function setState(state) {
    document.getElementById('myWidget').langevitour.setState(state);
}
</script>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

As an example we will use measurements of penguins made at Palmer Station, Antarctica. 

Since the measurements are in different units, we supply an appropriate scaling.

```{r}
library(langevitour)
library(palmerpenguins)

completePenguins <- na.omit(penguins[,c(1,3,4,5,6)])
completePenguins

scale <- apply(completePenguins[,-1], 2, sd)*4

langevitour(
    completePenguins[,-1], completePenguins$species, 
    scale=scale, pointSize=2, elementId="myWidget")
```

<br>

**Things to try:**

<div style="display: grid; grid-template-columns: 50% 50%; column-gap: 1em; row-gap: 1em; margin: 0 0 2em 4em;">
  <div>Drag labels onto the plot.</div>
  <div><button onclick="setState({labelPos:{'Adelie':[-0.6553191489361703,0.7595744680851064],'Chinstrap':[-0.5957446808510638,-0.7723404255319148],'Gentoo':[0.7829787234042553,0.06595744680851068]}})">Example</button> <button onclick="setState({labelPos:{}})">Clear</button></div>
  
  <div>Turn on a guide.</div>
  <div><button onclick="setState({pointRepulsionType:'pca'})">PCA</button> <button onclick="setState({guideType:'none'})">None</button></div>
  
  <div>Experiment with "damping" and "heat" in the hamburger menu.</div>
  <div>
  <button onclick="setState({damping:2,heat:1})">Twitchy</button>
  <button onclick="setState({damping:-2,heat:0})">Smooth</button>
  <button onclick="setState({damping:0,heat:3})">Hot</button>
  <button onclick="setState({damping:0,heat:-3})">Cool</button>
  <button onclick="setState({damping:0,heat:0})">Default</button>
  </div>
  
  <div>When only three measurements are shown, your eyes will interpret the display as 3D. Hide measurements by unchecking the checkbox on their label.</div>
  <div>
  <button onclick="setState({labelInactive:['bill_length_mm']})">Hide bill length</button>
  <button onclick="setState({labelInactive:[]})">Show all</button>
  </div>
</div>

# Further examples

Some further examples have been omitted from the CRAN package due to their large size and problematic dependencies. They can be viewed on the langevitour website:

* [Further examples](https://logarithmic.net/langevitour/articles/examples.html)
* [RNA-Seq extended example](https://logarithmic.net/langevitour/articles/rnaseq.html)


